---
title: "ojsr-vignette"
author: "Gaston Becerra"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ojsr-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(knitr)
```

# Overview

The aim of this package is to aid you in crawling OJS archives, issues, articles, galleys, and search results, and retrieving/scraping meta-data from articles.

**Important Notes:**

- **ojsr functions rely on OJS routing conventions** (<https://docs.pkp.sfu.ca/dev/documentation/en/architecture-routes>), therefore, it is most likely to fail on OJS installments with customized routing. 
- **ojsr get_\*_from_\* functions do not scrap the url you provide!**. ojsr will parse the urls you input to generate the *right* url to scrap in each scenario (e.g., the issue ToC url, if you are scrapping articles from an issue).

## About OJS

(from the OJS documentation <https://pkp.sfu.ca/ojs/>, as of Jan.2020)

Open Journal Systems (OJS) is a journal management and publishing system that has been developed by the Public Knowledge Project through its federally funded efforts to expand and improve access to research.

OJS assists with every stage of the refereed publishing process, from submissions through to online publication and indexing. Through its management systems, its finely grained indexing of research, and the context it provides for research, OJS seeks to improve both the scholarly and public quality of refereed research.

OJS is open source software made freely available to journals worldwide for the purpose of making open access publishing a viable option for more journals, as open access can increase a journal’s readership as well as its contribution to the public good on a global scale (see PKP Publications).

## OJS v3 and OJS API

Since OJS v3.1+ <https://docs.pkp.sfu.ca/dev/api/ojs/3.1> a Rest API is provided. We're positive a better R interface should use that API instead of webscraping. 

So, why ojsr? According to <https://pkp.sfu.ca/ojs/ojs-usage/ojs-stats/>, as of 2019 (when v3.1+ was launched), OJS was being used by at least 10,000 journals worldwide. OJS is an excellent free publishing solution for institutions that probably could not publish otherwise, and presumably, can not afford to update constantly. ojsr aims to help crawling and retrieving info from OJS during this legacy period.





# Use cases

xxxxx


# Function reference

## parse_base_url: Parses urls against OJS routing conventions to retrieve the base url

`parse_base_url()` takes a vector of OJS urls and retrieves their base url, according to OJS routing conventions.

```{r}
mix_links <- c(
   'https://dspace.palermo.edu/ojs/index.php/psicodebate/issue/archive',
   'https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/article/view/2903'
)
base_url <- ojsr::parse_base_url(input_url = mix_links)
```

Result is in a dataframe of the same length of your input:

1. *input_url* - the url you provided
2. *output_url* - the base url parsed


## get_issues_from_archive: Scraps issues' urls from OJS issue archive

`get_issues_from_archive()` takes a vector of OJS urls and scraps the issues urls from the *issue archive* (e.g., <https://papiro.unizar.es/ojs/index.php/rc51-jos/issue/archive>).

You don't need to provide the actual url to issue archives. `get_issues_from_archive()` parses the url you provide to compose it. 
Then, looks for links containing "/issue/view" in the href. Links are post-processed to comply to OJS routing conventions before returning.

```{r}
journals <- c( 
  'https://dspace.palermo.edu/ojs/index.php/psicodebate/issue/archive', # issue archive
  'https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/article/view/2903' # article
)
issues <- ojsr::get_issues_from_archive(input_url = journals, verbose = TRUE)
```

Result is in a long-format dataframe (1 input_url may result in several rows, one for each output_url), containing:

1. *input_url* - the url you provided
2. *output_url* - the issues' url scrapped

## get_articles_from_issue: Scrap articles urls from the ToC of OJS issues

`get_articles_from_issue()` takes a vector of OJS (issue) urls and scraps the links to articles from the issues table of content (e.g., <https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/issue/view/319/showToc>).

You don't need to provide the actual url of the issues' ToC, but you must provide urls that include issue ID (articles urls do not include this info!). `get_articles_from_issue()` parses the url you provide to compose the ToC url. Then, looks for links containing "/article/view" in the href. Links are post-processed to comply to OJS routing conventions before returning.

```{r}
issues <- c( 
  'https://revistas.ucn.cl/index.php/saludysociedad/issue/view/65', # issue including ToC
  'https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/issue/view/31' # no ToC nor links
)
articles <- ojsr::get_articles_from_issue(input_url = issues, verbose = TRUE) 
```

Result is in a long-format dataframe (1 input_url may result in several rows, one for each output_url), containing:

1. *input_url* - the url you provided
2. *output_url* - the articles url scraped

## get_articles_from_search: Scraps OJS search results for a given criteria to retrieve articles' url

`get_articles_from_search()` takes a vector of OJS urls and a string for search criteria to compose search result urls, 
then scraps them to retrieve articles' urls.

You don't need to provide the actual url of the search result pages. `get_articles_from_search()` parses the url you provide to compose the search result page(s) url. If pagination is involved, necessary links are also included. Then, looks for links containing "/article/view" in the href. Links are post-processed to comply to OJS routing conventions before returning.

```{r}
journals <- c( 
  'https://revistapsicologia.uchile.cl/index.php/RDP/', 
  'https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/issue/current' 
)
criteria <- "psicologia"
articles_searched <- ojsr::get_articles_from_search(input_url = journals, 
  search_criteria = criteria, verbose = TRUE)
```

Result is in a long-format dataframe (1 input_url may result in several rows, one for each output_url), containing:

1. *input_url* - the url you provided
2. *output_url* - the article url



## get_galleys_from_article: Scrap galleys urls from OJS articles

*Galleys* are the final presentation version of the content of the articles. Most of the time, these include the full-content in pdf and other reading formats. Less often, they are supplementary files (tables, dataset) in different formats. 

`get_galleys_from_article()` takes a vector of OJS urls and scraps all the galleys urls from the *article view* (e.g., <https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/article/view/593>).

You may provide any article-level url (article abstract view, inline view, pdf direct download, etc.). `get_galleys_from_article()` parses the url you provide to compose the url of the article view. Then, looks for links containing "/article/view" in the href. Links are post-processed to comply to OJS routing conventions before returning (i.e., having a galley ID).

```{r}
articles <- c( 
  'https://revistapsicologia.uchile.cl/index.php/RDP/article/view/55657', # galleys pdf and mp3
  'https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/article/view/2137', # pdf
  'https://dspace.palermo.edu/ojs/index.php/psicodebate/article/view/516/311' # inline reader
)
galleys <- ojsr::get_galleys_from_article(input_url = articles,verbose = TRUE) 
```

Result is in a long-format dataframe (1 input_url may result in several rows, one for each output_url), containing:

1. *input_url* - the url you provided
2. *output_url* - the galleys url scraped
3. *format* - the format of the galley (e.g., pdf, xml)
4. *download_url* - the conventional url to force download of the galley. You may pass these to a download function of your own (e.g., <https://stackoverflow.com/questions/39246739/download-multiple-files-using-download-file-function>).



## get_html_meta_from_article: Scrap metadata from html of OJS articles

`get_html_meta_from_article()` takes a vector of OJS urls and scraps all the metadata written in the html from the *article view* (e.g., <https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/article/view/593>).

You may provide any article-level url (article abstract view, inline view, pdf direct download, etc.). `get_html_meta_from_article()` parses the url you provide to compose the url of the article view. Then, looks for \<meta\> tags in the \<head\> section of the html. **Important!** This may not only retrieve bibliographic metadata; any other "meta" property detailed on the html will be obtained (e.g., descriptions for propagation on social network, etc.).

```{r}
articles <- c( 
  'https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/article/view/2137', # article
  'https://dspace.palermo.edu/ojs/index.php/psicodebate/article/view/516/311' # xml galley
)
metadata <- ojsr::get_html_meta_from_article(articles, verbose = TRUE) 
```

Result is in a long-format dataframe (1 inputUrl may result in several rows, one for each outputUrl), containing:

1. *input_url* - the url you provided
2. *meta_data_name* - name of the property/metadata (e.g., "DC.Date.created" for the Date of creation) 
3. *meta_data_content* - the actual value of the metatag
4. *meta_data_scheme* - the standard in which the content is annotated
5. *meta_data_xmllang* - the language in which the metadata was entered



## get_oai_meta_from_article: Retrieve OAI records for OJS articles

An alternative to webscraping metadata from the html of article pages is to retrieve their OAI-PMH (Open Archives Initiative Protocol for 'Metadata' Harvesting) records <http://www.openarchives.org/OAI/openarchivesprotocol.html>

`get_oai_meta_from_article()` will try to access the OAI records within the OJS for any article (e.g., <https://fundacionmenteclara.org.ar/revista/index.php/RCA/oai/?verb=GetRecord&identifier=oai:ojs.fundacionmenteclara.org.ar:article/43&metadataPrefix=oai_dc>) for which you provided an url. 

```{r}
articles <- c(  
  'https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/article/view/2137', # article
  'https://dspace.palermo.edu/ojs/index.php/psicodebate/article/view/516/311' # xml galley
)
metadata_oai <- ojsr::get_oai_meta_from_article(input_url = articles, verbose = TRUE)
```

Result is in a long-format dataframe (1 inputUrl may result in several rows, one for each outputUrl), containing:

1. *input_url* - the url you provided
2. *meta_data_name* - name of the property/metadata (e.g., "DC.Date.created" for the Date of creation) 
3. *meta_data_content* - the actual value of the metatag
4. *meta_data_scheme* - **always returns NA** (included just for easier binding with `get_html_meta_from_article()` results)
5. *meta_data_xmllang* - **always returns NA** (included just for easier binding with `get_html_meta_from_article()` results)

**Note: This function is in a very preliminar stage**. If you are interested in working with OAI records, you may want to check Scott Chamberlain's **OAI package for R** <https://CRAN.R-project.org/package=oai>. If you only have the OJS home url, and would like to check all the article's OAI records at one shot, an interesting option is to parse it with `ojsr::parse_oai_url()` and passing the output_url to `oai::list_identifiers()`.


## parse_oai_url: Parses urls against OJS routing conventions to retrieve the OAI protocol url

`parse_oai_url()` takes a vector of OJS urls and retrieves their OAI entry url, according to OJS routing conventions.

```{r}
mix_links <- c(
   'https://dspace.palermo.edu/ojs/index.php/psicodebate/issue/archive',
   'https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/article/view/2903'
)
oai_url <- ojsr::parse_oai_url(input_url = mix_links)
```

Result is in a dataframe of the same length of your input:

1. *input_url* - the url you provided
2. *output_url* - the base url parsed
