---
title: "ojsr-vignette"
author: "Gaston Becerra"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ojsr-vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(knitr)
```


# Overview

The aim of this package is to aid you in several OJS tasks, such as: 

- parse OJS urls, according to routing conventions;
- crawling archives, issues, articles, galleys;
- retrieving/scraping meta-data from articles.

## Important! 

- Most of ojsr functions rely on **OJS routing/links conventions and default themes** (see "process_urls()" below), and therefore are most likely to fail when used on customized OJS installments (i.e., customized themes, routing improvements via .htaccess), and on OJS configuration with poor record keeping.
- Only **OJS v1+ and v2+** have been considered (see "OJS v3 and OJS API" below).

## About OJS

(from the OJS documentation <https://pkp.sfu.ca/ojs/>, as of Jan.2020)

Open Journal Systems (OJS) is a journal management and publishing system that has been developed by the Public Knowledge Project through its federally funded efforts to expand and improve access to research.

OJS assists with every stage of the refereed publishing process, from submissions through to online publication and indexing. Through its management systems, its finely grained indexing of research, and the context it provides for research, OJS seeks to improve both the scholarly and public quality of refereed research.

OJS is open source software made freely available to journals worldwide for the purpose of making open access publishing a viable option for more journals, as open access can increase a journal’s readership as well as its contribution to the public good on a global scale (see PKP Publications).

## OJS v3 and OJS API

Since OJS v3.1+ <https://docs.pkp.sfu.ca/dev/api/ojs/3.1> a Rest API is provided. We're positive a better R interface should use that API instead of scraping web themes and/or OAI records. 

So, why ojsr? According to <https://pkp.sfu.ca/ojs/ojs-usage/ojs-stats/>, as of 2019 (when v3.1+ was launched), OJS was being used by at least 10,000 journals worldwide. OJS is an excellent free publishing solution for institutions that could not publish otherwise. Presumably, most of them do not have the the technical nor financial capabilities to update constantly, so OJS v1+ and v2+ are expected to be around for some time... 



# Installation

```{r message=FALSE, warning=FALSE}
# let's start by loading the libraries
library(ojsr) 
library(tidyverse) # we'll use dplyr and tidy several times in the examples
```



# How to use ojsr?

## process_urls: Parse urls to check what type of OJS page they point at

Let's say you have a bunch of OJS specific article urls you'd like to scrap their meta-data, or a curated list of journals you'd like to crawl completely. However, not all of them may be poiting to the *right* type of OJS page for your purposes (e.g., articles urls you collected may be downloading pdf directly, so you won't be able to scrap meta-data; or the issues may be pointing at cover pages that do not include ToCs). 

`process_urls()` allows you to **parse a list of urls against OJS routing conventions** to extract the parameters invoked (e.g., article, issues, galley IDs) and generate the *conventional* url for specific purposes, such as scraping meta-data, or scraping the list of issues for the archive of issues page.

`process_urls()` is invoked from the `get_*_url_from_*` functions before scraping when `use_contentional_url = TRUE` (or default). When a *conventional* url is used for scraping, this is also returned in the resulting dataframe, so you could easily join tables (from scraping articles, issues, archives, etc.) , by creating a table of conventional urls with `process_urls()`.

`process_urls()` works with **routing conventions** (see <https://docs.pkp.sfu.ca/dev/documentation/en/architecture-routes>). The list of urls you input is not actually browsed!

```{r message=FALSE, warning=FALSE}
# urls pointing at different types of OJS pages
url_sample <- c( 'https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/article/view/2903',  # article view
          'http://fundacionmenteclara.org.ar/revista/index.php/RCA/issue/view/2018-Vol3-2', # issue (with ToC)
          'http://fundacionmenteclara.org.ar/revista/index.php/RCA/article/download/43/54', # pdf download
          'https://firstmonday.org/ojs/index.php/fm/article/view/9540', # article view
          'http://imed.pub/ojs/index.php/iam/article/view/1650', # article view
          'http://fundacionmenteclara.org.ar/revista/index.php/RCA/oai', # oai protocol
          'https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/issue/view/444', # issue (no ToC)
          'https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/search/search?query=complejidad' # search result
)

# we'll parse them, and store the result in a dataframes (`what_type_of_urls`)
what_type_of_urls <- ojsr::process_urls(url_sample)
```

The resulting dataframe contains the following columns:

1. *url* - the url you provided
2. *baseUrl* - base url of the OJS (you may use this to join and relate tables generated with `get_*_url_from_*` functions)
3. *issueId* - if an issue, its ID (string); otherwise NA
4. *articleId* - if an article, its ID (string); otherwise NA
5. *galleyId* - if a galley (full-content, supplementary materials, etc.), its ID (string); otherwise NA
6. *conventional_archive* - the *conventional* form of the url you should use to scrap the list of issues
7. *conventional_issue* - the *conventional* form of the url you should use to scrap the ToC from an issue (if issueID is present, otherwise "")
8. *conventional_article* - the *conventional* form of the url you should use to scrap metadata from an article (if articleID is present, otherwise "")
9. *conventional_oai* - the *conventional* form of the url you should use to access OAI
10. *conventional_search* - the *conventional* form of the url you should use to scrap search results



## get_issue_urls_from_archive: Retrieve issues urls from the archive of issues

`get_issue_urls_from_archive()` takes a vector of OJS urls pointing at the archive page of a journal (e.g., <https://papiro.unizar.es/ojs/index.php/rc51-jos/issue/archive>) and scraps them to retrieve the links to issues.

If you don't know the url for the archive page, you may provide any url for the journal, and keep `use_conventional_url = TRUE` (default): ojsr will parse the urls given with `process_urls()` to form the conventional url ($conventional_archive) for the archive page.

Once the archive page is read, ojrs will scrap for links containing "/issue/view" (`method='scrap_by_href_convention'`).

```{r}
socPsy_urls <- c( # argentinian social psychology journals
  'https://dspace.palermo.edu/ojs/index.php/psicodebate/issue/archive', # points at the archive of issues
  'https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/article/view/2903' # points at an article; ojsr will use process_url() to form the right link to scrap
)
issues <- ojsr::get_issue_urls_from_archive(url = socPsy_urls, use_conventional_url = TRUE, verbose = TRUE)
```

Result is in a long-format dataframe, containing:

1. *archive* - the url you provided
2. *baseUrl* - if use_conventional_url = TRUE, the baseUrl from ojs::process_url; otherwise, ""
3. *issue* - the issues url



## get_article_urls_from_search_results: Retrieve articles urls from issue pages

`get_article_urls_from_search_results()` takes a vector of OJS urls pointing at the issues of a journal (e.g., <https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/issue/view/319/showToc>) and scraps them to retrieve the links to articles.

Issue pages in OJS may or may not show a cover page, and show or not show the table of contents (this is where the links to the article reside). Usually, you can force the display of the ToC by adding "/ShowToC" to the url. This is all taken care of if you keep `use_conventional_url = TRUE` (default): ojsr will parse the urls given with `process_urls()` to form the conventional url ($conventional_issue) for the issue page.

Once the issue page is read, ojrs can scrap for links using different criteria:

- links containing the "/article/view" convention, and then filtering only those with no "pdf" or "file" classes (usually pointing to a galley, such as, full-content, secondary reading formats, and/or supplementary files) (`method='scrap_by_href_convention_no_classes'`, default).
- links containing the "/article/view" convention, without further filtering (`method='scrap_by_href_convention'`). If the issue points both to the articles and their galleys, these may result in more than one link to an article.

```{r}
socPsy_issues_url <- c( #  initial issues from a few social psychology journals
  'https://revistas.ucn.cl/index.php/saludysociedad/issue/view/65', # includes ToC
  'https://dspace.palermo.edu/ojs/index.php/psicodebate/issue/view/Psicodebate%201', # includes ToC
  'https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/issue/view/31' # does not include ToC, nor links to articles
)
articles <- ojsr::get_article_urls_from_issue(socPsy_issues_url, use_conventional_url = TRUE, verbose = TRUE) 
```

Result is in a long-format dataframe, containing:

1. *issue* - the url you provided
2. *baseUrl* - if use_conventional_url = TRUE, the baseUrl from ojs::process_url; otherwise, ""
3. *articles* - the articles url




## get_galley_urls_from_article: Retrieve articles urls from article pages

`get_galley_urls_from_article()` takes a vector of OJS urls pointing at the articles of a journal (e.g., <https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/article/view/593>) and scraps them to retrieve the links to galleys.

*Galleys* are the final presentation version of the content of the articles. Most of the time, these include the full-content in pdf and other reading formats. Less often, they are supplementary files (tables, dataset) in different formats. 

Galleys are usually linked from other type of pages too, other than the view of the articles, like some issues ToC, or other galleys inline readers. However, the safest place to search for them is in the view of articles. If you provide an url with an article ID (it could be a view of the article, another galley, etc.), and if you keep `use_conventional_url = TRUE` (default), ojsr will parse the urls given with `process_urls()` to form the conventional url ($conventional_article) for the article view page.

Once the article page is read, ojrs will scrap for links containing classes "file", "download" or "obj_galley_link" (`method='scrap_by_class_convention'`, default).

```{r}
socPsy_articles <- c( # 3 articles on social psychology, specifically social representations theory
  'https://revistapsicologia.uchile.cl/index.php/RDP/article/view/55657', # 2 galleys: pdf and mp3
  'https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/article/view/2137', # 1 galley: pdf
  'https://dspace.palermo.edu/ojs/index.php/psicodebate/article/view/516/311' # url pointing to inline reader. also 2 galleys: pdf and xml
)
galleys <- ojsr::get_galley_urls_from_article(socPsy_articles, use_conventional_url = TRUE, verbose = TRUE) 
```

Result is in a long-format dataframe, containing:

1. *article* - the url you provided
2. *baseUrl* - if use_conventional_url = TRUE, the baseUrl from ojs::process_url; otherwise, ""
3. *galley* - the galley url linked
4. *format* - the format of the galley (e.g., pdf, xml)
5. *force* - the conventional url to force download of the galley

You may filter some of these, and then pass the "force" column to a download function of your own (e.g., <https://stackoverflow.com/questions/39246739/download-multiple-files-using-download-file-function>).




## get_metadata_from_article: Retrieve metadata from article pages

`get_metadata_from_article()` takes a vector of OJS urls pointing at the articles of a journal (e.g., <https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/article/view/593>) and scraps their metadata, stored in the <head> of the article html.

Metadata is usually present in the head of the article html, on the html that shows a galley. However, since some galley formats force the download of a file, the safes place to search for them is in the view of the article. If you provide an url with an article ID (it could be a view of the article, another galley, etc.), and if you keep `use_conventional_url = TRUE` (default), ojsr will parse the urls given with `process_urls()` to form the conventional url ($conventional_article) to scrap metadata in the head of the html.

Once the html is read, ojrs will scrap for <meta> tags in the <head> section of the article's html (`method='scrap_meta_in_head'`, default). **Important!** This will not only retrieve bibliographic meta-data; any other meta property detailed on the article's html will also be captured (e.g., descriptions for propagation on social network, etc.).

```{r}
socPsy_articles <- c( # 3 articles on social psychology, specifically social representations theory
  'https://revistapsicologia.uchile.cl/index.php/RDP/article/view/55657', # url pointing to the article page
  'https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/article/view/2137', # url pointing to the article page
  'https://dspace.palermo.edu/ojs/index.php/psicodebate/article/view/516/311' # url pointing a particular galley (xml)
)
metadata <- ojsr::get_metadata_from_article(socPsy_articles, use_conventional_url = TRUE, verbose = TRUE) 
```
Result is in a long-format dataframe, containing:

1. *article* - the url you provided
2. *baseUrl* - if use_conventional_url = TRUE, the baseUrl from ojs::process_url; otherwise, ""
3. *meta_data_name* - name of the property/metadata (e.g., "DC.Date.created" for the Date of creation) 
4. *meta_data_content* - the actual value of the metatag
5. *meta_data_scheme* - the standard in which the content is annotated
6. *meta_data_xmllang* - the language in which the metadata was entered

You may analize this dataframe to explore bibliometric properties of a collection of articles.




## get_oai_metadata_from_article: Retrieve OAI records from article url

An alternative to webscraping metadata from the html of article pages is to retrieve their OAI-PMH (Open Archives Initiative Protocol for 'Metadata' Harvesting) records <http://www.openarchives.org/OAI/openarchivesprotocol.html>

`get_oai_metadata_from_article()` will try to access the OAI records for any OJS article page url. Since several parameters and url need to the formed, this function always invokes `ojsr::process_urls()`.

**Note:** If you are interested in working with OAI records, you may want to check Scott Chamberlain's **OAI package for R** <https://CRAN.R-project.org/package=oai>. If you only have the OJS home url, and would like to check all the article's OAI records at one shot, an interesting option is to parse it with `ojsr::process_urls()` and passing the OAI url ($conventional_oai) to `oai::list_identifiers()`.

```{r}
socPsy_articles <- c( # 3 articles on social psychology, specifically social representations theory
  'https://revistapsicologia.uchile.cl/index.php/RDP/article/view/55657', # url pointing to the article page
  'https://publicaciones.sociales.uba.ar/index.php/psicologiasocial/article/view/2137', # url pointing to the article page
  'https://dspace.palermo.edu/ojs/index.php/psicodebate/article/view/516/311' # url pointing a particular galley (xml)
)
metadata_oai <- ojsr::get_oai_metadata_from_article(url = socPsy_articles, verbose = TRUE)
```

Result is in a long-format dataframe, containing:

1. *article* - the url you provided
2. *baseUrl* - if use_conventional_url = TRUE, the baseUrl from ojs::process_url; otherwise, ""
3. *meta_data_name* - name of the property/metadata (e.g., "DC.Date.created" for the Date of creation) 
4. *meta_data_content* - the actual value of the metatag
